---
title: "Untitled"
author: "Jared Flater"
date: "10/8/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
```{r}
library(tidyverse)
library(phyloseq)
library(phylosmith)
```
# Overall Praire vs. Crops at both sites
```{r}
WORARM <- readRDS("data/all_samples_merged_STRIPS.RDS")
soil <- subset_samples(WORARM, matrix == c("soil"))
pcoa_phyloseq(soil, c('soil_type'), circle = TRUE)
```

# Worle soil
 
```{r setup, include=FALSE}
worle <- readRDS("data/Worle_curated.RDS")
```

This data now includes bacterial communities from soil samples from the WOR rainfall experiment. 

```{r}
soil <- subset_samples(worle, matrix == c("soil"))
head(sample_data(soil))
```

# Are bacterial communities in soil from strip different from crop? 

Look at all baseline samples, before manure was applied. Depth is a strong driver of clustering. Seperate into two datasets by depth, S1 and S2.
```{r}
s1 <- subset_samples(soil, depth == 1)
s2 <- subset_samples(soil, depth == 2)
```

Visualize if samples from crop are different from prairie. Select baseline samples and plot PCoA ordinations.  
```{r}
s1plot <- s1 %>%
  subset_samples(sample_day == c("Baseline")) %>%
  pcoa_phyloseq(c('soil_type'), circle = TRUE)
s2plot <- s2 %>%
  subset_samples(sample_day == c("Baseline")) %>%
  pcoa_phyloseq(c('soil_type'), circle = TRUE)
s1plot
s2plot
```
Rarefied and mintax 10
```{r}
d1rmintax10 <- subset_samples(soil, depth == 1 & sample_day == c("Baseline")) %>%
  rarefy_even_depth(sample.size = 10000, rngseed = 1231, replace = T) %>%
  filter_taxa(function(x) sum(x) >= 10, T) %>%
  pcoa_phyloseq(c('soil_type'), circle = TRUE)

d2rmintax10 <- subset_samples(soil, depth == 2 & sample_day == c("Baseline")) %>%
  rarefy_even_depth(sample.size = 10000, rngseed = 1231, replace = T) %>%
  filter_taxa(function(x) sum(x) >= 10, T) %>%
  pcoa_phyloseq(c('soil_type'), circle = TRUE)
d1rmintax10
d2rmintax10
```

