---
title: "Strips soil manure associated taxa"
author: "Jared Flater"
date: "Spring 2020"
output:
  html_document:
    df_print: paged
    fig_width: 8.5
    fig_height: 11
    fig_caption: true
    toc: true
    toc_float: true
    number_sections: false
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r More Libraries}
library(phyloseq)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(kableExtra)
library(venn)
library(phylosmith)
library(phylosmith)
```

# Load RDS

```{r Read RDS}
worle <- readRDS("../data/worle_tax_are_seqs.RDS")
arm <- readRDS("../data/armstrong_tax_are_seqs.RDS")
worle 
arm
```

# Make names for environmental and experimental variables consistent

```{r Sample Data}
arm_data1 <- data.frame(sample_data(arm))

worle_data1 <- data.frame(sample_data(worle))
```
```{r arm table}
kable(arm_data1) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "600px")
```
```{r worle table}
kable(worle_data1) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "600px")
```
## Change WORLE
```{r Change WORLE}
worle_data <- worle_data1  %>%
  dplyr::select(unique_id = id, experiment = PI, matrix, treatment, plot, sample_day = day, depth, in_plot_location, block) %>%
  mutate(experiment = "Worle_Rainfall") %>%
  mutate(strip = ifelse(treatment %in% c("no_manure_strip", "manured_strip"), "Y", "N")) %>%
  mutate(manure_treatment = ifelse(treatment %in% c("manured_control", "manured_strip"), "Y", "N")) %>%
  mutate(soil_type = ifelse(in_plot_location %in% c("s1", "s2", "s3", "s4"), "crop",
                            ifelse(in_plot_location %in% c("s6", "s7", "s8", "s9"), "strip", "border"))) %>%
  mutate(sample_day = ifelse(sample_day == 'b', "Baseline",
                             ifelse(sample_day == "t0", "T000",
                                    ifelse(sample_day == "t2", "T002",
                                           ifelse(sample_day == "t14", "T014", 
                                                  ifelse(sample_day == "t21", "T021", "T042"
                                                 )))))) %>%
  mutate(depth = ifelse(depth == "d1", "1", "2"))
worle_data$in_plot_location[worle_data$in_plot_location == "S6"] <- "s6"
kable(worle_data) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "600px")
```

## Change Armstrong

```{r Armstrong Data change}
arm_data <- arm_data1 %>%
  mutate(in_plot_location = sample_num) %>%
  mutate(treatment = ifelse(treatment == 'ACS', "no_manure_strip",
                             ifelse(treatment == "ACM", "manured_control",
                                    ifelse(treatment == "ACSM", "manured_strip", "NA"
                                          )))) 
arm_data$sample_day[arm_data$sample_day == "TB"] <- "Baseline"
arm_data$in_plot_location <- paste0('s', arm_data$in_plot_location) # WORLE has s before in_plot_location

kable(arm_data) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "600px")
```
```{r Re-make sample data and save as RDS}
rownames(arm_data) <- arm_data$unique_id
rownames(worle_data) <- worle_data$unique_id
sample_data(arm) <- arm_data
nrow(sample_data(arm))
nrow(arm_data)
sample_data(worle) <- worle_data
nrow(sample_data(worle))
nrow(worle_data)
saveRDS(object = worle, file = "../data/Worle_curated.RDS")
saveRDS(object = arm, file = "../data/Arm_curated.RDS")
rm(arm_data, arm_data1, worle_data, worle_data1)
```

```{r Merge Phsyeqs}
worle <- readRDS("../data/Worle_curated.RDS")
worle
arm <- readRDS("../data/Arm_curated.RDS")
arm

STRIPS <- merge_phyloseq(worle, arm)
STRIPS
saveRDS(STRIPS, file = "../data/all_samples_merged_STRIPS.RDS")
```
```{r}
### Scale color for phyla, consistent phyla colors across plots
#Function to make all phyla colors consistent

phylalist <- data.frame(tax_table(STRIPS),row.names = NULL) %>%
  select(Phylum) %>%
  unique() 
phylalist$Phylum <- as.character(phylalist$Phylum)
phylalist$Phylum[is.na(phylalist$Phylum)] <- "Unclassified"

# this package will generate a pallette based on number and desired colors
library(colorspace)
colors <- sequential_hcl(n_distinct(phylalist), palette = "viridis") %>%
  sample() %>%
  setNames(phylalist$Phylum)
rm(STRIPS)
```

```{r plot armstrong manure}
arm_manure <- subset_samples(arm, matrix == c("manure"))
sample_names(arm_manure)
data.frame(sample_sums(arm_manure)) %>%
  rownames_to_column(var = "Sample") %>%
  as_tibble() %>%
  ggplot(aes(x = reorder(Sample, sample_sums.arm_manure.), y = sample_sums.arm_manure.)) + 
  geom_bar(stat = "identity")

arm_manure_mintax1 <- arm_manure %>%
  filter_taxa(function(x) sum(x) >= 1, T) 
rm(arm_manure)
data.frame(sample_sums(arm_manure_mintax1)) %>%
  rownames_to_column(var = "Sample") %>%
  as_tibble() %>% 
  ggplot(aes(x = reorder(Sample, sample_sums.arm_manure_mintax1.), y = sample_sums.arm_manure_mintax1.)) + 
  geom_bar(stat = "identity")

arm_manure_asvs <- taxa_names(arm_manure_mintax1)

armphylo <- phylogeny_profile(arm_manure_mintax1, classification = "Phylum") 
 
armphylo + scale_color_manual(aesthetics = "fill", values = colors)
```

```{r plot worle manure}
worle_manure <- subset_samples(worle, matrix == c("manure"))
sample_names(worle_manure)
data.frame(sample_sums(worle_manure)) %>%
  rownames_to_column(var = "Sample") %>%
  as_tibble() %>%
  ggplot(aes(x = reorder(Sample, sample_sums.worle_manure.), y = sample_sums.worle_manure.)) + 
  geom_bar(stat = "identity")

worle_manure_mintax1 <- worle_manure %>%
  filter_taxa(function(x) sum(x) >= 1, T) 
rm(worle_manure)
data.frame(sample_sums(worle_manure_mintax1)) %>%
  rownames_to_column(var = "Sample") %>%
  as_tibble() %>% 
  ggplot(aes(x = reorder(Sample, sample_sums.worle_manure_mintax1.), y = sample_sums.worle_manure_mintax1.)) + 
  geom_bar(stat = "identity")

worle_manure_asvs <- taxa_names(worle_manure_mintax1)

worlephylo <- phylogeny_profile(worle_manure_mintax1, classification = "Phylum") 

worlephylo + scale_color_manual(aesthetics = "fill", values = colors)
```

```{r}
arm_soil_mintax1 <- subset_samples(arm, matrix == c("soil") & manure_treatment == c("N")) %>%
  filter_taxa(function(x) sum(x) >= 1, T)
nsamples(arm_soil_mintax1)
kable(data.frame(sample_data(arm_soil_mintax1), row.names = NULL)) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "600px")

arm_soil_asvs <- taxa_names(arm_soil_mintax1)
```

```{r}
worle_soil_mintax1 <- subset_samples(worle, matrix == c("soil") & manure_treatment == c("N")) %>%
  filter_taxa(function(x) sum(x) >= 1, T)
nsamples(arm_soil_mintax1)
kable(data.frame(sample_data(worle_soil_mintax1), row.names = NULL)) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "600px")

worle_soil_asvs <- taxa_names(worle_soil_mintax1)
```

# Venn diagrams of manure ASVs

## Armstrong Venn

List of manure associated ASVs not detected in soil samples, left side of the venn

```{r}
arm_vvv_diag <- venn(list("Manure_ASVs" = arm_manure_asvs, "Soil_ASVs" = arm_soil_asvs))
arm_manure_persitors <- attr(arm_vvv_diag, "intersections")$Manure_ASVs
```

## Worle venn

```{r}
worle_vvv_diag <- venn(list("Manure_ASVs" = worle_manure_asvs, "Soil_ASVs" = worle_soil_asvs))
worle_manure_persitors <- attr(worle_vvv_diag, "intersections")$Manure_ASVs
```

```{r}
arm.soil.phy <- subset_samples(arm, matrix == "soil") %>%
  filter_taxa(function(x) sum(x) >= 1, T) %>% 
  transform_sample_counts(function(x) x / sum(x))
unique(arm.soil.phy@sam_data$treatment)

arm.persistors.phy <- prune_taxa(arm_manure_persitors, arm.soil.phy)

ntaxa(arm.persistors.phy)
kable(data.frame(sample_data(arm.persistors.phy), row.names = NULL)) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "600px")
```

```{r}
worle.soil.phy <- subset_samples(worle, matrix == "soil") %>%
  filter_taxa(function(x) sum(x) >= 1, T) %>% 
  transform_sample_counts(function(x) x / sum(x))

worle.persistors.phy <- prune_taxa(worle_manure_persitors, worle.soil.phy)

ntaxa(worle.persistors.phy)
kable(data.frame(sample_data(worle.persistors.phy), row.names = NULL)) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%", height = "600px")
```
# Plots

## Armstrong plot

```{r}
arm.melted <- psmelt(arm.persistors.phy)
unique(arm.melted$treatment)
arm.g <- ggplot(arm.melted, aes(x = in_plot_location, y = Abundance))
arm.plot <- arm.g + 
  geom_bar(aes(fill = Phylum), stat = "identity") +
  facet_grid(sample_day~treatment)
arm.plot + scale_color_manual(aesthetics = "fill", values = colors)
```

## WORLE plot

```{r}
worle.melted <- psmelt(worle.persistors.phy)

worle.g <- ggplot(worle.melted, aes(x = in_plot_location, y = Abundance))

worle.plot <- worle.g + 
  geom_bar(aes(fill = Phylum), stat = "identity") +
  facet_grid(sample_day~treatment)
worle.plot + scale_color_manual(aesthetics = "fill", values = colors)
```

It appears that Armstrong has a significantly higher portion of manure associated bacteria, we should investigate the validity of this...


```{r}

```



































































